A catalog of wide binaries and comoving star pairs from \textsl{Gaia DR1} and low-resolution spectroscopy

### 2017-06-29 paper/comoving-followup.tex
more intro

\begin{abstract}
% Context
Widely-separated, coeval stars are powerful tools for testing star-formation
model predictions, improving stellar atmosphere models, and inferring the
mass distribution of the Milky Way.
They vastly outnumber the Galactic open clusters, but are still expected to
share a common origin and chemistry.
The kinematic properties of bound wide binaries and their unbound complements
serve as potentiometers for the smooth tidal field and perturbation spectrum in
the Milky Way.
Thus far, a few hundred genuine comoving star pairs with separations as large
as $8~\pc$ have been discovered in the solar neighborhood.
% Aims
In previous work, we have identified $\sim 10,000$ candidate comoving star pairs
using only astrometric data from the Tycho-Gaia Astrometric Solution (TGAS)
catalog, a part of \gaia\ \DR{1}.
However, without radial velocity measurements for the vast majority of the
pairs, the contamination rate at large separations is expected to be
considerable.
% Methods
Here, we present results from our own radial-velocity survey of a sample of the
candidate pairs to validate wide-separation comoving star pairs.
% Results
Of the XX comoving pairs observed, we find YY ...
% Conclusions
Gaia DR2 -> how many do we expect?
\end{abstract}

@@

The majority of stars are thought to form in dynamically cold, chemically
homogeneous clusters, at least relative to the kinematic and chemical
dispersions observed amongst typical stars in the Galactic disk.
Because star formation proceeds rapidly (e.g., \citealt{Bovy:2016}) and the
progenitor gas cloud disperses quickly (cite), it is typically assumed that
stars formed in the same birth cluster have the same age and initial chemistry
(e.g., cite).
These birth clusters are typically born un- or loosely-bound and therefore
quickly succumb to the tidal field of the Milky Way; obvious dynamical
associations between stars from the same cluster are then erased as the
constituent stars mix into the general population of disk stars.
However, the stars do preserve their initial associations in age, chemical
abundance patterns, and dynamical invariants (e.g., orbital actions).
It has therefore been long presumed that ``chemical tagging'' and chemodynamical
models of the Galactic disk will allow for inferences about the formation and
evolution of the Milky Way.
That is, the chemical environment of a birth cluster is likely imprinted on its
stellar children so that they may be re-associated with precisely measured
abundance patterns (of $\gtrsim 10$ species).

@@

Are these properties of open clusters universal, or limited to the few clusters
presently studied? And do the chemical properties of open clusters depend on the
star formation rate or efficiency, which may evolve with time?
Ideal tests of the chemical homogeneity assumptions would use many more open
clusters with varied cluster masses, and member stars that span a large range in
masses and evolutionary stages.
However, because they disrupt quickly, the present-day open cluster population
is biased towards younger or more massive clusters.

An alternate tracer of the chemical properties mentioned above are
widely-separated, coeval pairs or multiplets of stars.
Summary of wide binary literature

- Also useful for stellar models: should have same age, chemistry, but need to
  test this
    - Chemical homogeneity tested with few open clusters - wide binaries another
      test
    - Need sample of confirmed widely-separated binaries, follow-up to get
      detailed abundances
    - Need way to verify that widely separated stars are physically associated,
      via kinematics
- Tests of chemical enrichment due to planetary systems

In this work, we present a catalog of genuine comoving pairs.
We expect high-resolution follow-up to study chemistry...

@@

Outline:
- Stars form in clusters, chemically homogeneous at some level
- Clusters generally disrupt slowly through a process that preserves initial
  clustering in conserved quantities but destroys phase coherence
- Most significant process is probably tidal heating
- Dissolved clusters are therefore useful dynamical tracers of mass distribution
- In Galactic halo, globular clusters
- In disk, hope is to use chemical signatures to connect stars, then can use
  dynamics to reconstruct birth clusters, infer mass distribution / evolution
- Clusters have advantage of many stars per cluster, but fewer clusters
- Wide binary stars are little potentiometers themselves - only N=2, but are far
  more numerous as a population
- Long-term goal is to study the separation distribution of wide-binaries as a
  function of age and Galactic orbit distribution (disk vs. halo)
    - Is there a separation truncation?
    - Can that be described just from Galactic tides, or need perturbers?

- Some bullshit about Gaia, TGAS, etc., released as a part of \gaia\ \DR{1}\

\cite{Gaia-Collaboration:2016,Gaia-Collaboration:2016a}

comoving not wide binary (don't know if they are bound)

Widely-separated comoving stars are expected to form from the dissolution of
birth star clusters

>>>>

\begin{abstract}
% Context
Widely-separated, coeval stars are powerful tools for testing star-formation
model predictions, improving stellar atmosphere models, and inferring the
mass distribution of the Milky Way.
They vastly outnumber the Galactic open clusters, but are still expected to
share a common origin and chemistry.
The kinematic properties of bound wide binaries and their unbound complements
serve as potentiometers for the smooth tidal field and perturbation spectrum in
the Milky Way.
Thus far, a few hundred genuine comoving star pairs with separations as large
as $8~\pc$ have been discovered in the solar neighborhood.
% Aims
In previous work, we have identified $\sim 10,000$ candidate comoving star pairs
using only astrometric data from the Tycho-Gaia Astrometric Solution (TGAS)
catalog, a part of \gaia\ \DR{1}.
However, without radial velocity measurements for the vast majority of the
pairs, the contamination rate at large separations is expected to be
considerable.
% Methods
Here, we present results from our own radial-velocity survey of a sample of the
candidate pairs to validate wide-separation comoving star pairs.
% Results
Of the XX comoving pairs observed, we find YY ...
% Conclusions
Gaia DR2 -> how many do we expect? Can we say anything? Out to what distance and
separation for a typical MS star can we determine they are bound assuming
high-res spectra to get very precise RVs?
\end{abstract}

@@

The majority of stars are thought to form in dynamically cold, chemically
homogeneous clusters, at least relative to the kinematic and chemical
dispersions observed amongst typical stars in the Galactic disk.
Because star formation proceeds rapidly (e.g., \citealt{Bovy:2016}) and the
progenitor gas cloud disperses quickly (cite), it is typically assumed that
stars formed in the same birth cluster have the same initial chemistry
(e.g., cite + \citealt{Feng:2014}) and same ages.
These birth clusters are typically born un- or loosely-bound and therefore
quickly succumb to the tidal field of the Milky Way; obvious dynamical
associations between stars from the same cluster are then erased as the
constituent stars mix into the general population of disk stars.
However, the stars do preserve their initial associations in age, chemical
abundance patterns, and dynamical invariants (e.g., orbital actions).
It has therefore been long presumed that ``chemical tagging'' and chemodynamical
models of the Galactic disk will allow for inferences about the formation and
evolution of the Milky Way.
That is, the chemical environment of a birth cluster is likely imprinted on its
stellar children so that they may be re-associated with precisely measured
abundance patterns (of $\gtrsim 10$ species).

@@

Are these properties of open clusters universal, or limited to the few clusters
presently studied?
Do the chemical properties of open clusters depend on the star formation rate or
efficiency, which may evolve with time?
Ideal tests of the chemical homogeneity assumptions would use many more open
clusters with varied cluster masses and member stars that span a large range in
masses and evolutionary stages.
\todo{Gaia will help \gaia\ (\cite{Gaia-Collaboration:2016,Gaia-Collaboration:2016a})}
However, because they disrupt quickly, the present-day open cluster population
is biased towards younger or more massive clusters.

An alternate tracer of the chemical properties mentioned above are
widely-separated ($>XX~\pc$), coeval pairs or multiplets of stars.
These tracers include both bound and unbound wide-separation stellar systems; we
refer to these systems collectively as ``comoving stars'' because, at large
separations, with present-day and near-future velocity uncertainties, it is
typically not possible to determine whether these systems are in fact bound.
Comoving stars are either wide binaries, disrupted wide binaries, or remnants of
dissolving or disrupted open clusters.
In any of these cases, it is likely that the stars formed from the same region
of a birth cloud, yet are separated enough that there has not been any
cross-pollution in the event of strong stellar winds or stellar evolution.

For chemical studies, open clusters have the advantage of having many stars per
cluster, but there are many fewer clusters than comoving stars.
Thanks to the \project{Hipparcos} mission (cite) and large radial-velocity
surveys (cite rave, etc.), several hundred such systems are already known
(\citealt{Shaya:2011}).
In addition, over ten-thousand \emph{candidate} comoving stars have been
discovered from astrometric data alone, using data from both the
\project{Hipparcos} mission (\citealt{hipparcos?}) and the \gaia\ mission
(\citealt{Oh:2017,Andrews:2017}).
- Need sample of confirmed widely-separated binaries, follow-up to get
  detailed abundances
- Need way to verify that widely separated stars are physically associated,
  via kinematics

Once chemical homogeneity understood amongst stars of same spectral type, can be used to calibrate stellar models across the HR diagram: should have same age, chemistry, etc. Also tests of chemical enrichment due to planetary systems

Comoving stars are also sensitive potentiometers of the Galactic mass
distribution.
- Long-term goal is to study the separation distribution of wide-binaries as a
  function of age and Galactic orbit distribution (disk vs. halo)
    - Is there a separation truncation?
    - Can that be described just from Galactic tides, or need perturbers?

In this work, we present a catalog of genuine comoving star pairs discovered
through radial velocity follow-up of candidate comoving stars discovered from
their astrometry alone (\citealt{Oh:2017}).
We expect high-resolution follow-up to study chemistry...

@@
